---
title: "OpenML"
author: "The OpenML R Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OpenML}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r init, include=FALSE}
# library("knitr")
# opts_chunk$set(cache = TRUE)
library("OpenML")
setOMLConfig(apikey = "c1994bdb7ecb3c6f3c8f3b35f4b47f1f")
```


# Listing {#listing}
In this stage, we want to list basic information about the various OpenML objects such as data sets, tasks, flows, runs, run results, evaluation measures or task types.
See the [OpenML introduction](http://openml.org/guide) for an overview on and explanations of the different objects.

For each of these objects we have a function to query the information beginning with `listOML`. All of these functions return a `data.frame`, even when the result has only one column.

First, load the package:
```{r results="hide"}
library("OpenML")
setOMLConfig(verbosity = 0) # switch off status output
```

### List data sets
To browse the OpenML data base for appropriate data sets, you can use `listOMLDataSets()` in order to get basic data characteristics (number of features/instances/classes/missing values etc.) for each data set.
By default, `listOMLDataSets()` returns only data sets that have an active status on OpenML.
If you need data sets that are either `"in_preparation"` or `"deactivated"`, you can change the `status` parameter:
```{r}
datasets = listOMLDataSets()  # returns active data sets
names(datasets)
head(datasets[, 1:6])

inactive.data = listOMLDataSets(status = "deactivated")  # returns deactivated data sets
head(inactive.data[, 1:6])
```

To find a specific data set, one can now query the resulting `datasets` object. Suppose we want
to find the `iris` data set.
```{r}
subset(datasets, name == "iris")
```
As one can see there are two data sets called `iris`.
We want to use the original data set with three classes.
It has the data set ID `did` = `r subset(datasets, name == "iris" & NumberOfClasses == 3)$did`.
One can also look at the data set on the OpenML web page http://openml.org/d/`r subset(datasets, name == "iris" & NumberOfClasses == 3)$did`.

### List tasks
Each OpenML task is a bundle of a data set, a target feature, a (performance) estimation procedure (e.g., 10-fold CV), data splits for this estimation procedure and, finally, one or more (performance) evaluation measures.
Every task has a type, e.g., `"Supervised Classification"` or `"Supervised Regression"`.
To list tasks one can use
```{r}
tasks = listOMLTasks()
names(tasks)
head(tasks[, 1:5])
```
For some data sets, there may be more than one task available at the OpenML server.
For example, you can look for `"Supervised Classification"` tasks that are available for a specific data set as follows:
```{r}
# subset tasks to "Supervised Classification" for the iris data (did == 61)
head(subset(tasks, task.type == "Supervised Classification" & did == 61L)[, 1:5])
```

### List flows
A flow is the definition and implementation of a specific algorithm workflow or script.
I.e., a flow is essentially the code that implements the algorithm.
```{r}
flows = listOMLFlows()
names(flows)
flows[56:63, 1:2]
```

### List runs and run results
A run is a combination of a setup (**??? FIXME??? Setup is not explained (ans why should one use it if it cannot be queried?**) and a task.
The results are stored as a run result.
Both, runs and run results can be listed.
Here one has additional arguments to subset the result.
For example one can search for a specific `task.id`, `setup.id` or `implementation.id`.
To list all runs for [task 59](http://www.openml.org/t/59) (supervised classification on iris) one can use
```{r}
runs = listOMLRuns(task.id = 59L)  # must be restricted to a task, setup and/or implementation ID
head(runs)

# FIXME: This does not work in the vignette?
# runresults = listOMLRunEvaluations(task.id = 61)  # a task ID must be supplied
# colnames(runresults)
```

### List evaluation measures and task types
To list further objects such as evaluation measures and task types one can simply use the respective functions.
```{r}
listOMLEvaluationMeasures()
listOMLTaskTypes()
```
