---
title: "OpenML"
author: "The OpenML R Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OpenML}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r init, include=FALSE}
# library("knitr")
# opts_chunk$set(cache = TRUE)
library("OpenML")
setOMLConfig(apikey = "c1994bdb7ecb3c6f3c8f3b35f4b47f1f")
```

# Downloading {#download}
Users can download data, tasks, flows and runs from the OpenML server.
The package provides special representations for each object which will be discussed here.

## Download an OpenML data set only
To directly download a data set, e.g., when you want to run a few preliminary experiments, one can use the function `getOMLDataSet`.
The function accepts a data set ID as input and returns the corresponding `OMLDataSet`:
```{r}
iris.data = getOMLDataSet(did = 61L)  # the iris data set has the data set ID 61
iris.data
```

## Download an OpenML task
The following call returns an OpenML task object for a supervised classification task on the iris data:
```{r}
task = getOMLTask(task.id = 59L)
task
```
The corresponding `"OMLDataSet"` object can be accessed by
```{r}
task$input$data.set
```
A special print function gives the basic information on the data set.
To extract the data itself one can use
```{r}
iris.data = task$input$data.set$data
head(iris.data)
```


## Download an OpenML flow
You can download a flow by specifying the `flow.id` parameter in the `getOMLFlow` function:
```{r}
flow = getOMLFlow(flow.id = 2700L)
flow
```

## Download an OpenML run
To download the results of one run including all server and user computed metrics, you have to know the corresponding run ID.
These IDs can be extracted from the `runs` object, which was created in the previous section.
Here we use the first run of task 59, which has the `run.id` 234.
You can download a single OpenML run with the `getOMLRun` function:
```{r}
run = getOMLRun(run.id = 525534L)  # see ?OMLRun for each slot of the OMLRun object
```
There are some slots which are of major interest for the `OMLRun` object.
A list containing the parameter settings can be obtained by `run$parameter.setting`:
```{r}
run$parameter.setting  ## retrieve first parameter setting
```
(**??? FIXME: WHAT DOES THIS REPRESENT???**)
The first three parameters refer to the RNG settings, e.g. the seed, which will help you to reproduce the run.
It the flow that created this run has hyperparameters that are different from the default values of the corresponding learner, they are also shown, otherwise the default hyperparameters are used.

All data that served as input for the run, including the URL to the data is stored in
```{r}
run$input.data
```
(**??? FIXME: STRANGE???**)

To retrieve predictions of an uploaded run, you can use:
```{r}
head(run$predictions)
```

