Listing
=======

In this stage, we want to list basic information about the various OpenML objects such as data sets, tasks, flows, runs, run results, evaluation measures or task types. See the [OpenML introduction](http://openml.org/guide) for an overview on and explanations of the different objects.

For each of these objects we have a function to query the information beginning with `listOML`. All of these functions return a `data.frame`, even when the result has only one column.

First, load the package:
```{r results="hide"}
library("OpenML")
setOMLConfig(verbosity = 0) # switch off status output
```

### List data sets
To browse the OpenML data base for appropriate data sets, you can use `listOMLDataSets()`
in order to get basic data characteristics (number of features/instances/classes/missing values etc.)
for each data set. By default, `listOMLDataSets()` returns only data sets that have an active
status on OpenML. If you need data sets that are either `"in_preparation"` or `"deactivated"`,
you can change the `status` parameter:

```{r}
datasets = listOMLDataSets()  # returns active data sets
datasets[1:3, 3:6]

inactive.data = listOMLDataSets(status = "deactivated")  # returns deactivated data sets
inactive.data[1:3, 3:6]
```

To find a specific data set, one can now query the resulting `datasets` object. Suppose we want
to find the `iris` data set.

```{r}
subset(datasets, name == "iris")[, 1:6]
```

As one can see there are two data sets called `iris`. We want to use the original data set
with three classes. It has the data set ID `did` = `r subset(datasets, name == "iris" & NumberOfClasses == 3)$did`.
One can also look at the data set on the OpenML web page
http://openml.org/d/`r subset(datasets, name == "iris" & NumberOfClasses == 3)$did`.

### List tasks
Each OpenML task is a bundle of a data set, a target feature, a (performance) estimation procedure (e.g.,
10-fold CV), data splits for this estimation procedure and, finally, one or more (performance) evaluation measures.
Every task has a type, e.g., `"Supervised Classification"` or `"Supervised Regression"`. To list tasks one can use
```{r}
tasks = listOMLTasks()
tasks[1:6, 1:5]
```

For some data sets, there may be more than one task available at the OpenML server.
For example, you can look for `"Supervised Classification"` tasks that are available
for a specific data set as follows:

```{r}
tasks = listOMLTasks()
# subset tasks to "Supervised Classification" for the iris data (did == 61)
subset(tasks, task_type == "Supervised Classification" & did == 61L)[, c(1, 6)]
```
### List flows

A flow is the definition and implementation of a specific algorithm workflow or script.
I.e., a flow is essentially the code that implements the algorithm.

```{r}
flows = listOMLFlows()
flows[56:63, 1:2]
```

### List setups

A setup is the combination of a flow with specific parameter settings.

**FIXME: Add this once it is up and running**

### List runs and run results

A run is a combination of a setup and a task. The results are stored as a run result.
Both, runs and run results can be listed. Here one has additional arguments to subset
the result. For example one can search for a specific `task.id`, `setup.id` or `implementation.id`.
To list all runs for [task 59](http://www.openml.org/t/59).

```{r eval = FALSE}
runs = listOMLRuns(task.id = 59L)  # must be restricted to a task, setup and/or implementation ID
head(runs)

runresults = listOMLRunResults(task.id = 59L)  # a task ID must be supplied
colnames(runresults)
```

### List evaluation measures and task types
To list further objects such as evaluation measures and task types one can simply use
the respective functions.

```{r}
measures = listOMLEvaluationMeasures()
measures[1:7, , drop = FALSE]

tasktypes = listOMLTaskTypes()
tasktypes
```


----------------------------------------------------------------------------------------------------
Jump to:
- [Introduction](1-Introduction.md)
- [Configuration](2-Configuration.md)
- Listing
- [Downloading](4-Downloading.md)
- [Running models on tasks](5-Running.md)
- [Uploading](6-Uploading.md)
- [Example workflow with mlr](7-Example-workflow-with-mlr.md)
