Stage 0 - Listing
=================

In this stage, we want to list basic information about the various OpenML objects such as data sets, tasks, flows, runs, run results, evaluation measures or task types. See the [OpenML introduction](http://openml.org/guide) for an overview on and explanations of the different objects. 

For each of these objects we have a function to query the information beginning with `listOML`. All of these functions return a `data.frame`, even when the result has only one column.

First, load the package:
```{r results="hide"}
library("OpenML")
setOMLConfig(verbosity = 0) # switch of status output
```

### List data sets 
To browse the OpenML data base for appropriate data sets, you can use `listOMLDataSets()`
in order to get basic data characteristics (number of features/instances/classes/missing values etc.)
for each data set. By default, `listOMLDataSets()` returns only data sets that have an active
status on OpenML. If you need data sets that are either `"in_preparation"` or `"deactivated"`,
you can change the `status` parameter:

```{r}
datasets = listOMLDataSets()  # returns active data sets
datasets[1:3, 3:6]

inactive.data = listOMLDataSets(status = "deactivated")  # returns deactivated data sets
inactive.data[1:3, 3:6]
```

To find a specific data set, one can now query the resulting `datasets` object. Let's use
the `iris` data set.

```{r}
subset(datasets, name == "iris")[, 1:6]
```

As one can see that there are two data sets called `iris`. We use the original three class
data set. It has the data set ID `did` = `r subset(datasets, name == "iris" & NumberOfClasses == 3)$did`. 
One can also look at the data set on the OpenML web page http://openml.org/d/`r subset(datasets, name == "iris" & NumberOfClasses == 3)$did`.

### List tasks
To list tasks one can use
```{r}
tasks = listOMLTasks()
tasks[1:6, 1:5]
```

For some data sets, there may be more than one task available at the OpenML server. 
For example, you can look for `"Supervised Classification"` tasks that are available 
for a specific data set as follows:

```{r}
tasks = listOMLTasks()
# subset tasks to "Supervised Classification" for the iris data (did == 61)
subset(tasks, task_type == "Supervised Classification" & did == 61L)[, c(1, 6)]
```

### List runs and run results

**FIXME: How is listOMLRuns**
Similarly, you can also list runs and run results for a specific task, setup or implementation:

```{r eval = FALSE}
runs = listOMLRuns(task.id = 1L)  # must be restricted to a task, setup and/or implementation ID
head(runs)

runresults = listOMLRunResults(task.id = 1L)  # a task ID must be supplied
colnames(runresults)
```

### List flows, evaluation measures and task types
To list further objects such as flows, evaluation measures and task types one can simply use 
the respective functions.

```{r}
flows = listOMLFlows()
flows[1:7, 1:2]

measures = listOMLEvaluationMeasures()
measures[1:7, , drop = FALSE]

tasktypes = listOMLTaskTypes()
tasktypes
```


----------------------------------------------------------------------------------------------------
Jump to:
[Introduction](1-Introduction.md)
[Configuration](2-Configuration.md)
Stage 0 - Listing
[Stage 1 - Downloading](4-Stage-1-Downloading.md)
[Stage 2 - Running models on tasks](5-Stage-2-Running.md)
[Stage 3 - Uploading](6-Stage-3-Uploading.md)
[8 Example workflow with mlr](8-Example-workflow-with-mlr.md)
