Browse the database
===================

To browse the data base for appropriate data sets, you can use the so called data qualities. Data qualities contain basic data characteristics (number of features/instances/classes/missing values
etc.) as well as values of more abstract meta learning features. 

### Get data qualities
With the function `getDataQualities` you can obtain all data qualities of all stored data sets. 
By default, only the basic data characteristics are retrieved:
```{r}
dq = getDataQualities()
```

If you want to retrieve not only the basic data qualities but also meta learning features, 
please use the argument "set":
```{r eval = FALSE}
dq = getDataQualities(set = "all")
```

Now, you have a data.frame with some data qualities. You can use them to find data sets that meet
certain conditions. Let's assume we wanted to work on data sets that have only numeric features and
not a single missing value. Therefore, we simply subset the data.frame:
```{r}
filtered.dq = subset(dq, NumberOfNumericFeatures == NumberOfFeatures & NumberOfMissingValues == 0)
head(filtered.dq[, c("did", "dataset", "version")])
```

These are the names, versions and data set IDs of all appropriate data sets that are stored on the 
OpenML server. Now we could download one or more of these data sets directly via
`downloadOpenMLDataSet` as seen in [section 2](2-Download-a-task.md). Another option is to find
tasks that work with these data sets and download one or more of them:
```{r, eval = FALSE}
task.ids = getTasksForDataSet(filtered.dq$did)
```

### Make an arbitrary SQL-query
The freest way to browse the OpenML database is by SQL-queries. The function `runSQLQuery` is 
an interface for any arbitrary SQL-query. The query is passed as a string to the function as 
you can see in the following example:
```{r eval = FALSE}
runSQLQuery(query = "SELECT id FROM implementation WHERE name = 'classif.rpart'")
```

Note that most users should not need this function.

----------------------------------------------------------------------------------------------------
Jump to:   
[1 Introduction](1-Introduction.md)  
[2 Download a task](2-Download-a-task.md)  
[3 Upload an implementation](3-Upload-an-implementation.md)  
[4 Upload predictions](4-Upload-predictions.md)  
[5 Download performance measures](5-Download-performance-measures.md)  
6 Browse the database
[7 Example workflow with mlr](7-Example-workflow-with-mlr.md)