Uploading
=========

### Upload an implementation using mlr

A flows is an implementation of single algorithms or scripts. To create an implementation, we can use the `mlr` package. Each `mlr` learner can be considered as an implementation of a flow that can be uploaded to the server with the function `uploadOMLFlow`.
If the implementation has already been uploaded to the server, we get a message that the flow already exists and the `implementation.id` is returned from the function. Otherwise, the not existing flow is uploaded and a new `implementation.id` is assigned to it.

```{r eval = TRUE, warning = FALSE, message = TRUE}
library(mlr)
lrn = makeLearner("classif.randomForest")
implementation.id = uploadOMLFlow(lrn, session.hash = session.hash)
implementation.id
```

### Upload an implementation without using mlr

In the previous section, we explained how to create an `OMLFlow` manually and created the object `oml.flow`, which reflects implementation description object. Before you can upload this flow to the server, you have to write an R-script containing the algorithm you want to use as flow. Let's assume you have done this and have a string `sourcefile` containing the path to your R-script. Your flow can now be uploaded as follows:

```{r eval = FALSE}
oml.flow.id = uploadOMLFlow(oml.flow, sourcefile = sourcefile)
```

### Upload an OpenML run to the server

Runs that have been created using `mlr` should have an `implementation.id` so that you can upload the run using:

```{r eval = TRUE, warning = FALSE, message = TRUE}
run.id = uploadOMLRun(run.mlr)
```

A manually created run might not contain an `implementation.id` so that you can upload the run and manually assign a `implementation.id`:

```{r eval = FALSE}
run.id = uploadOMLRun(run, implementation.id = oml.flow.id)
```

----------------------------------------------------------------------------------------------------
Jump to:
- [Introduction](1-Introduction.md)
- [Configuration](2-Configuration.md)
- [Listing](3-Listing.md)
- [Downloading](4-Downloading.md)
- [Running models on tasks](5-Running.md)
- Uploading
- [Example workflow with mlr](7-Example-workflow-with-mlr.md)