Uploading
=========

```{r init2, include=FALSE}
library("OpenML")
## temporarily set API key to read only key
setOMLConfig(apikey = "c1994bdb7ecb3c6f3c8f3b35f4b47f1f")
setOMLConfig(verbosity = 0) # switch off status output
```

```{r recreate_resuls, include=FALSE}
task = getOMLTask(task.id = 59L)
iris.data = task$input$data.set$data
library(mlr)
lrn = makeLearner("classif.rpart")
run.mlr = runTaskMlr(task, lrn)
run.mlr
```

(**??? Fixme: We need to make the results from section 5 available here. Is there an easy and good way to source all chunks from sec. 5 here but do not include the output? http://yihui.name/knitr/demo/child/ seems not to work.**)

### Upload an implementation using mlr

(**??? Fixme: Shoudln't we explain first, why we need to upload the implementation before uploading the results? And isn't that needed only once?**)

(**??? Fixme: Shouldn't we better call the `implementation.id` `flow.id`?**)

A flows is an implementation of single algorithms or scripts. To create an implementation, we can use the `mlr` package. Each `mlr` learner can be considered as an implementation of a flow that can be uploaded to the server with the function `uploadOMLFlow`.
If the implementation has already been uploaded to the server, we get a message that the flow already exists and the `implementation.id` is returned from the function. Otherwise, the not existing flow is uploaded and a new `implementation.id` is assigned to it.

```{r eval = TRUE, warning = FALSE, message = TRUE}
library(mlr)
lrn = makeLearner("classif.randomForest")
implementation.id = uploadOMLFlow(lrn)
implementation.id
```
(**??? Fixme: What happens if a flow already exists on the server? And what happens here if we use the read-only API key? The tutorial states that we get a message that the flow already exists. This is not the case!**)

### Upload an implementation without using mlr

In the previous section, we explained how to create an `OMLFlow` manually and created the object `oml.flow`, which reflects implementation description object. Before you can upload this flow to the server, you have to write an R-script containing the algorithm you want to use as flow. Let's assume you have done this and have a string `sourcefile` containing the path to your R-script. Your flow can now be uploaded as follows:

```{r eval = FALSE}
oml.flow.id = uploadOMLFlow(oml.flow, sourcefile = sourcefile)
oml.flow.id
```
(**??? Fixme: What happens here? The code runs without error despite the fact that sourcefile doesn't exist.**)


### Upload an OpenML run to the server

Runs that have been created using `mlr` should have an `implementation.id` so that you can upload the run using:

```{r eval = FALSE, warning = FALSE, message = TRUE}
run.id = uploadOMLRun(run.mlr)
```
(**??? Fixme: Cannot be executed with read-only API key. I still would like to execute this in order to check the code...**)

(**??? Fixme: Not working anymore. How to proceed? How do we get the flow.id in the runs?**)
~~A manually created run might not contain an `implementation.id` so that you can upload the run and manually assign a `implementation.id`:~~
```{r eval = FALSE}
run.id = uploadOMLRun(run, implementation.id = oml.flow.id)
```

----------------------------------------------------------------------------------------------------
Jump to:
- [Introduction](1-Introduction.md)
- [Configuration](2-Configuration.md)
- [Listing](3-Listing.md)
- [Downloading](4-Downloading.md)
- [Running models on tasks](5-Running.md)
- Uploading
- [Example workflow with mlr](7-Example-workflow-with-mlr.md)